Temporal Dynamics of Normalization Reweighting

### 2023-07-31 manuscript.Rmd
Testing build from processed individual data

processdata <- 1  # this flag determines the amount of processing, with 4 levels:
# 0 - do no processing, generate the pdf using existing versions of all figures
# 1 - generate figures using the processed group data (requires 0.4GB of storage)
# 2 - download processed individual data and analyse (requires 11.5GB of storage)
# 3 - download all raw data and analyse (requires 143GB of storage)

@@

Following the suggestion of a reviewer, we conducted an alternative fixed-phase analysis, where the signal in each 1 second epoch was multiplied by a sine wave of appropriate frequency and phase instead of taking the Fourier transform. The results were similar to our main analysis, and can be viewed in the Figures subdirectory of the project code repository ([https://github.com/bakerdh/normreweight/tree/main/Figures files with the suffix 'FP'](https://github.com/bakerdh/normreweight/tree/main/Figures)). We also conducted simulations (also available in the code repository) to confirm that our analysis methods were not distorting the estimates of the suppression timecourse. In brief, although the 1 second sliding time window does blur the signals in time, these effects are largely negated by calculating the suppression ratio because the blur cancels out across the numerator and denominator. Overall these simulations give us confidence in the accuracy of our estimates of suppression dynamics.

@@

Next we repeated the experiment on 20 participants using a 248-channel whole-head cryogenic MEG system. Half of the participants had a diagnosis of autism, and the remainder were age- and gender-matched controls. Source localisation using a linearly constrained minimum variance (LCMV) beamformer algorithm [@VanVeen1997] showed strong localisation of steady-state signals at the occipital pole (see Figure \@ref(fig:MEGdata)a), and in the Fourier domain (Figure \@ref(fig:MEGdata)b). Responses from the most responsive V1 vertex showed a similar timecourse to those of the EEG experiments at both frequencies (Figure \@ref(fig:MEGdata)c,d), and showed increasing suppression during the first few seconds of stimulus presentation (Figure \@ref(fig:MEGdata)e,f). The normalization reweighting effect was again clearest at 5Hz, especially for the dichoptic condition (red curve in Figure \@ref(fig:MEGdata)e), which increased until 2.5s. This confirms that the reweighting effects can occur as early as primary visual cortex, consistent with findings from neurophysiology [@Aschner2018]. However the data are more variable than for our EEG experiments, and had fewer significant clusters, perhaps owing to a power reduction caused by the smaller sample size for this dataset and greater heterogeneity across frequency. When pooling effects over frequency and condition, the overall effect size (*d* = `r round(MEGeffectsize,digits=2)`) was near zero.

>>>>

processdata <- 2  # this flag determines the amount of processing, with 4 levels:
# 0 - do no processing, generate the pdf using existing versions of all figures
# 1 - generate figures using the processed group data (requires 0.4GB of storage)
# 2 - download processed individual data and analyse (requires 11.5GB of storage)
# 3 - download all raw data and analyse (requires 143GB of storage)

@@

Following the suggestion of a reviewer, we conducted an alternative fixed-phase analysis, where the signal in each 1 second epoch was multiplied by a sine wave of appropriate frequency and phase instead of taking the Fourier transform. The results were similar to our main analysis, and can be viewed in the Figures subdirectory of the project code repository ([https://github.com/bakerdh/normreweight/tree/main/Figures - see files with the suffix 'FP'](https://github.com/bakerdh/normreweight/tree/main/Figures)). We also conducted simulations (also available in the code repository) to confirm that our analysis methods were not distorting the estimates of the suppression timecourse. In brief, although the 1 second sliding time window does blur the signals in time, these effects are largely negated by calculating the suppression ratio because the blur cancels out across the numerator and denominator. Overall these simulations give us confidence in the accuracy of our estimates of suppression dynamics.

@@

Next we repeated the experiment on 20 participants using a 248-channel whole-head cryogenic MEG system. Half of the participants had a diagnosis of autism, and the remainder were age- and gender-matched controls. Source localisation using a linearly constrained minimum variance (LCMV) beamformer algorithm [@VanVeen1997] showed strong localisation of steady-state signals at the occipital pole (see Figure \@ref(fig:MEGdata)a), and in the Fourier domain (Figure \@ref(fig:MEGdata)b). Responses from the most responsive V1 vertex showed a similar timecourse to those of the EEG experiments at both frequencies (Figure \@ref(fig:MEGdata)c,d), and showed increasing suppression during the first few seconds of stimulus presentation (Figure \@ref(fig:MEGdata)e,f). The normalization reweighting effect was again clearest at 5Hz, especially for the dichoptic condition (red curve in Figure \@ref(fig:MEGdata)e), which increased until 2.5s. This confirms that the reweighting effects can occur as early as primary visual cortex, consistent with findings from neurophysiology [@Aschner2018]. However the data are more variable than for our EEG experiments, and had fewer significant clusters, perhaps owing to a power reduction caused by the smaller sample size for this dataset and/or greater heterogeneity across frequency. When pooling effects over frequency and condition, the overall effect size (*d* = `r round(MEGeffectsize,digits=2)`) was near zero.
